---
title: "Reproducible Research"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Data
##The data for this assignment can be downloaded from the course web site:
###Dataset: Activity monitoring data [52K]

###The variables included in this dataset are:
####steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
####date: The date on which the measurement was taken in YYYY-MM-DD format
####interval: Identifier for the 5-minute interval in which measurement was taken
####The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

##Loading and preprocessing the data
###1.Load the data (i.e. read.csv())

```{r}
getwd()
setwd("F:/R/repdata%2Fdata%2Factivity")
data <- read.csv("activity.csv")
head(data)
```
###2.Process/transform the data (if necessary) into a format suitable for your analysis

##What is mean total number of steps taken per day?
###1.Calculate the total number of steps taken per day
```{r}
steps_per_day <- aggregate(steps ~ date, data, sum)
```

###2.If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day
```{r}
hist(steps_per_day$steps, main = paste("Total Steps Taken Each Day"),  xlab="Number of Steps")
```

###3.Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean_total_steps <- mean(steps_per_day$steps, na.rm = TRUE)
mean_total_steps
median_total_steps <- median(steps_per_day$steps, na.rm = TRUE)
median_total_steps
```

##What is the average daily activity pattern?
###1.Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
steps_by_interval <- aggregate(steps ~ interval, data, mean)

plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="5-minute Interval", ylab="Number of Steps", main="Average Number of Steps per Day by Interval")
```


###2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of step
```{r}
max_interval <- steps_by_interval[which.max(steps_by_interval$steps),1]
max_interval
```

##Imputing missing values
###Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

###1.Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(data$steps))
```
###2.Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
#### mean of the 5-minute interval 

###3.Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
data_new <- transform(data, steps = ifelse(is.na(data$steps), steps_by_interval$steps[match(data$interval, steps_by_interval$interval)], data$steps))

sum(is.na(data_new))
```
###4.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
imp_steps_per_day <- aggregate(steps ~ date, data_new, sum)
hist(imp_steps_per_day$steps, main = paste("Total Steps Each Day"), col="green", xlab="Number of Steps")

#Create Histogram to show difference. 
hist(steps_per_day$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("green", "blue"), lwd=10)

imp_mean <- mean(imp_steps_per_day$steps)
imp_mean
imp_median <- median(imp_steps_per_day$steps)
imp_median

mean_diff <- imp_mean - mean_total_steps
mean_diff
med_diff <- imp_median - median_total_steps
med_diff

total_diff <- sum(imp_steps_per_day$steps) - sum(steps_per_day$steps)
total_diff
```

##Are there differences in activity patterns between weekdays and weekends?
###For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

###1.Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

data_new$Weektype = as.factor(ifelse(is.element(weekdays(as.Date(data_new$date)),weekdays), "Weekday", "Weekend"))

head(data_new)

imp_steps_by_interval <- aggregate(steps ~ interval + Weektype, data_new, mean)
```

###2.Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r}
library(lattice)
xyplot(imp_steps_by_interval$steps ~ imp_steps_by_interval$interval|imp_steps_by_interval$Weektype, main="Average Steps per Day by Interval",xlab="5-minute Interval", ylab="Average Steps",layout=c(1,2), type="l")

```
